---
title: "Evaluation"
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(dplyr)
library(portalr)
source("tools/forecast_tools.R")

models_of_interest = c('AutoArima','Ensemble','ESSS','nbGARCH','pevGARCH')

species_of_interest = c('BA','DM','DO','PP','OT','NA','total')
species_names = portalr::FullPath('PortalData/Rodents/Portal_rodent_species.csv', '~') %>%
  read.csv(stringsAsFactors=FALSE, na.strings = '') %>%
  select(species = speciescode, full_species_name = scientificname)

#add a total entry
species_names = species_names %>%
  add_row(species='total', full_species_name='Total Rodents')

new_moon_file = portalr::FullPath('PortalData/Rodents/moon_dates.csv', '~')
new_moons = read.csv(new_moon_file)
new_moons$censusdate = as.Date(new_moons$censusdate)

# Setup rodent observations
species_abundance = portalr::abundance("repo", shape='flat', level='treatment', length="longterm") %>%
  filter(treatment=="control") %>%
  rename(actual = abundance, level = treatment) %>%
  mutate(currency='abundance') %>%
  left_join(new_moons, by='period') %>%
  filter(species %in% species_of_interest)

total_abundance = portalr::abundance("repo", shape='flat', level='treatment', length="longterm") %>%
  filter(treatment=="control") %>%
  group_by(period) %>%
  summarise(actual=sum(abundance)) %>%
  ungroup() %>%
  mutate(level='treatment',currency='abundance',species='total') %>%
  left_join(new_moons, by='period')

observation_data = species_abundance %>%
  bind_rows(total_abundance)

```

## How have the models done recently?

This graph shows the forecast error of some recent forecasts for the most recent sampling dates. This is shown for all model types in the columns and several common rodent species (and total rodents) in the rows. These evaluations are for the Control plots where no manipulation is done. 
```{r recent_eval, echo=FALSE, message=TRUE, warning=FALSE,, fig.width=15, fig.height=15}
###############################################
#Get all the recent forecasts
forecast_data = compile_forecasts(use_hindcasts = FALSE) %>%
  filter(model %in% models_of_interest)

# Round so that large decimal place differences do not cause
# similar forecasts to be seen as distinct. 
forecast_data$estimate = round(forecast_data$estimate, 3)
forecast_data$LowerPI = round(forecast_data$LowerPI, 3)
forecast_data$UpperPI = round(forecast_data$UpperPI, 3)

# Keep 1 forecast per unique initial_newmoon. When there are multiple forecasts
# using the same initial_newmoon, use the first one.
forecast_dates_to_keep = forecast_data %>%
  select(date, initial_newmoon) %>%
  distinct() %>%
  group_by(initial_newmoon) %>%
  filter(date == min(date)) %>%
  ungroup() %>%
  mutate(keep='yes') 

forecast_data = forecast_data %>%
  left_join(forecast_dates_to_keep, by=c('date','initial_newmoon')) %>%
  filter(keep=='yes') %>%
  select(-keep)

##################################
#Combine forecasts and oservations
forecast_errors = forecast_data %>%
  left_join(observation_data, by=c('species','level','currency','newmoonnumber')) %>% 
  mutate(rmse = sqrt((estimate - actual)^2))

# Sometimes there are model runs on the same day and with all the same info,
# this gets rid of those
forecast_errors = forecast_errors %>%
  distinct()

# Drop any entries that don't have an observation
# (ie. a rainy sample period)
forecast_errors = forecast_errors %>%
  filter(!is.na(actual))

# Filter to common species and apply the full name
forecast_errors = forecast_errors %>%
  filter(species %in% species_of_interest) %>%
  left_join(species_names, by='species') %>%
  select(-species) %>%
  rename(species=full_species_name)

# Add a newline to the longer species names
forecast_errors$species = with(forecast_errors, ifelse(nchar(species)>15, stringr::str_replace(species,' ','\n'), species))

#############################################################
# Displaying the data from many forecasts in the past can create a mess of a plot. see #218.
# So here three forecasts are chosen from 1, 6, and 12 newmoons into the past. 
# Sometimes forecasts data is missing for a particular initial_newmoon, the recursive function here
# attempts to find another one from the same time pluts/minus 2 month period. This ensures 3 forecasts
# are always shown.

# If there is no data for initial_newmoon 500, this will check nearby dates in this order: 499,501,498,402
update_initial_newmoon = function(potential_initial_newmoon, attempt=1){
  if(attempt==6){
    return(NA)
  }
  lead_time_data = forecast_errors %>%
    filter(initial_newmoon == potential_initial_newmoon)
  
  # Alternate steps of -1, +2, -3, +4
  next_try_step = ifelse(attempt%%2 == 0, attempt, attempt * -1)
  
  # no data to show? then try one month back
  if(nrow(lead_time_data)==0){
    #print(paste('Try ',attempt, ', Nothing with ',potential_initial_newmoon,' trying ',potential_initial_newmoon+next_try_step))
    return(update_initial_newmoon(potential_initial_newmoon+next_try_step, attempt+1))
  } else {
    return(potential_initial_newmoon)
  }
  
}
# Adjust this to add more than 3 forecast dates, or change the lead times
lead_times_to_show = c(1,6,12)
most_recent_observed_newmoon = max(new_moons$newmoonnumber)
initial_newmoons_to_keep = most_recent_observed_newmoon - lead_times_to_show

initial_newmoons_to_keep = sapply(initial_newmoons_to_keep, update_initial_newmoon)

forecast_errors = forecast_errors %>%
  filter(initial_newmoon %in% initial_newmoons_to_keep)

#####################################
ggplot(forecast_errors, aes(x=censusdate, y=rmse, group=as.character(date), color=as.character(date))) +
  geom_point(size=4)+
  geom_line(size=1.5) +
  scale_color_brewer(palette='Dark2') +
  scale_x_date(breaks = unique(forecast_errors$censusdate)) +
  facet_grid(species~model, scales = 'free_y') + 
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=0.99, debug = FALSE),
        axis.text = element_text(size=12),
        axis.title = element_text(size=20),
        strip.text = element_text(size=12),
        legend.text = element_text(size=15),
        legend.title = element_text(size=18)) +
  labs(x='Sample Date',y='RMSE', color='Forecast Issue\nDate')

```



## How have the models done historically?

These graphs show errors as a function of lead time. The lead time is the number of months into the future that forecast is made. The error values are an average of all forecast errors using observations since 2010. Note that this currently uses hindcasts of the prior observations, and is also only for the Control plots.

**RMSE**: Root mean square error, this is a metric used to evaluate the point estimate of a forecast.  
**Coverage**: This is the percentage of observations which fell within the 95% confidence interval of a forecast. Ideally this would be equal to 0.95. If it's higher than 0.95 the forecasts intervals are too wide, if it's lower then the forecast intervals are too narrow.


```{r hindcast_eval, echo=FALSE, message=TRUE, warning=FALSE,, fig.width=9, fig.height=15}

#Get the all the forecasts made during observation period
forecast_data = compile_forecasts(use_hindcasts = TRUE)
forecast_errors = calculate_forecast_error(observation_data, forecast_data, error_metric = 'RMSE') %>%
  filter(error_value < 200) %>% #Drop RMSE greater than this because it throws off all the graphs
  bind_rows(calculate_forecast_error(observation_data, forecast_data, error_metric = 'coverage'))

forecast_errors = forecast_errors %>%
  left_join(species_names, by='species')

ggplot(forecast_errors, aes(x=lead_time, y=error_value, group=model, color=model)) +
  geom_point()+
  geom_line() +
  geom_hline(yintercept = 0.95) +
  labs(x='Lead Time (New Moons)') +
  facet_wrap(full_species_name~error_metric, scales = 'free_y', ncol=2) + 
  theme_bw() +
  labs(y = "Error Value", colour = "Model")


```
