---
title: "Portal Forecast"
---

Welcome to Portal Forecasting! This is a website run by the [Weecology](http://weecology.org/) group, a group of interdisciplinary ecologists at the [University of Florida](http://www.wec.ufl.edu/). On this website, you'll find information about our ongoing efforts to forecast ecological systems from time series. Specifically, we are using a times series of rodent abundances from [The Portal project](http://portal.weecology.org/), a long-term experimental monitoring project in desert ecology. Enjoy! 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
'%>%' <- magrittr::'%>%'
source("tools/forecast_tools.R")
```

## Total Abundance Forecast

This is the forecast for next month's sampling of rodents at Portal. The black line indicates the historic total rodent abundance. The blue line indicates the forecasted amount over the next 12 months, along with confidence intervals. This forecast is from the Ensemble model. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
obs_data = portalr::abundance(path = "repo", clean=FALSE, level='treatment', plots="longterm") %>%
  dplyr::filter(treatment=="control")
obs_data$total = rowSums(dplyr::select(obs_data, -c(period, treatment)))
new_moons = read.csv(
  "https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv") %>%
                   dplyr::mutate(censusdate = as.Date(censusdate),
                          year = lubridate::year(newmoondate),
                          month = lubridate::month(newmoondate))

obs_data_newmoon = dplyr::inner_join(obs_data, new_moons, by = c("period" = "period"))

for_data = compile_forecasts() %>%
  dplyr::filter(level == 'Controls', model == 'Ensemble', date == max(date))

# get dates for future new moons for plotting
future_moons = new_moons %>% portalr::get_future_moons() %>%    
                         dplyr::select(c("newmoonnumber","newmoondate"))
for_data = dplyr::inner_join(for_data, future_moons, by = c("newmoonnumber" = "newmoonnumber"))

forecast_viz(obs_data = obs_data_newmoon,
             obs_date_col_name = "censusdate",
             obs_val_col_name = "total",
             for_data = for_data,
             for_date_col_name = "newmoondate",
             for_val_col_name = "estimate",
             for_model_name = "Ensemble",
             for_lowerpi_col_name = "LowerPI",
             for_upperpi_col_name = "UpperPI",
             start_newmoon = 300,
             ylabel = 'Total Abundance')
```

## Species-Level Forecasts

Point forecasts and 90% confidence intervals for specific species at the site for the next sampling period. 

```{r speciesforecasts, echo=FALSE, message=FALSE, warning=FALSE}

sp_predictions = get_sp_predicts(for_data, 'Controls', 
            lead_time = length(unique(for_data$newmoonnumber)) - 11)
sp_predict = plot_species_forecast(sp_predictions,
            title=paste0(sp_predictions$forecast_date[2], ": Control plots"))

plot(sp_predict)
```

## Most Abundant Species

This figures are similar to the primary Total Abundance forecast above, but for the abundance of three common rodent species. 
```{r highabund_ts_forecasts, echo=FALSE, message=FALSE, warning=FALSE}
most_abund_sp = sp_predictions %>% 
  dplyr::filter(species != 'total') %>% 
  dplyr::arrange(desc(estimate)) %>% 
  head(3) %>% 
  dplyr::select(species)

# load in rodent species table to get scientific names to display on plots
species_table = read.csv(
  "https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/Portal_rodent_species.csv",
                          stringsAsFactors = F,na.strings = '')
most_abund_sp_names = dplyr::filter(species_table,speciescode %in% most_abund_sp$species) %>%      
                      dplyr::select(speciescode,scientificname)

for (n in seq(dim(most_abund_sp_names)[1])) {
  species_forecast = forecast_viz(obs_data = obs_data_newmoon,
               obs_date_col_name = "censusdate",
               obs_val_col_name = most_abund_sp_names$speciescode[n],
               for_data = for_data,
               for_date_col_name = "newmoondate",
               for_val_col_name = "estimate",
               for_model_name = "Ensemble",
               for_lowerpi_col_name = "LowerPI",
               for_upperpi_col_name = "UpperPI",
               start_newmoon = 300,
               ylabel = most_abund_sp_names$scientificname[n])
  plot(species_forecast)
}
```

